
```{r readingData, echo=FALSE, warning=FALSE, message=FALSE}
# read in the data - add the dataset's path to filePath
library(readxl)
dataset <- read_excel("data/Data.xlsx", sheet = "data1")
#incomeData <- read_excel("data/Data.xlsx", sheet = "real") #importing real income
source("code/data_clean.R")
log_data <- data_clean(dataset)

source("code/dataset_dummy_YR.R")
dsetYRdummy <- dataset_dummy_YR(log_data)
source("code/dataset_dummy_YR_lag_Ex.R")
dsetYRdummyLagEx <- dataset_dummy_YR_lag_Ex(dsetYRdummy, log_data)
vecm_data <- dsetYRdummy %>% 
  rename("log_REAL" = 3)

#source("code/interpolate.R") #interpolate real income to monthly intervals and clean and log income data
# monthly_income <- interpolate(incomeData)
# vecm_data <- log_data %>% mutate(monthly_income$log_REAL, .before=log_YR)
# colnames(vecm_data)[12] = "log_REAL"
```

# Introduction

This study examines the relationship between the legal and illegal tobacco markets in South Africa. Section \ref{data} discusses the data used and how it was cleaned. Section \ref{methodology} explains the methodology, where a VECM model is presented. The final section details discussion points (\ref{analysis}). The appendix contains the full model outputs. 


# The Illicit Tobacco Market

The health problems associated with tobacco consumption are significant and have been well-documented (\citealp *{tingum2020}:62). Worldwide, smoking is one of the leading causes of preventable deaths; and in South Africa 23% of the total deaths in 2018 were smoking-related (Statistics South Africa). Given the large and negative impact of smoking, the tobacco market has come under increasing scrutiny by governments and health organizations. A significant component of the tobacco market is the illicit cigarette trade, which the Financial Action Task Force (2012:7) defines as 

> “*…the supply, distribution and sale of smuggled genuine, counterfeit or cheap white tobacco products.*”


The illicit tobacco market gives rise to concern on three main fronts: public health, tax evasion, and criminal activity. From a public health perspective, illicit cigarettes make smoking more affordable and thereby increases tobacco access (\citealp* {vanderzee2019}:242). Cheaper cigarettes may induce non-smokers to smoke, increase the volume of cigarettes consumed by smokers, and decrease the likelihood that smokers will quit smoking (\citealp* {pechacek2018}:pagenumber; \citealp{vanderzee2019}:242).


According to the International Agency for Research on Cancer, tobacco use is more prevalent among low socio-economic groups, and the poor are more sensitive to cigarette prices (IARC, 2011:276). Consequently, the health problems linked to smoking are disproportionately higher among the poor, which results in a greater burden on the public healthcare sector (). In South Africa, the healthcare costs associated with smoking amounted to R14.48 billion in 2016, which accounted for 4.1% of health spending (). In addition to the healthcare costs, there are indirect costs linked to smoking such as the loss of productive lives and loss of productive days due to illness (\citealp* {boachie2020}:pg). In 2016, the total cost of smoking to South Africa was R42.32 billion (0.97% of GDP).


Evidently, the public health and economic costs of tobacco use are extensive. In an effort to reduce tobacco use, many countries have implemented tax and price policies on tobacco products (Chaloupka, Straif, Leon). For a middle-income country, South Africa was considered to be at the forefront of tobacco control policies, including excise duties, for many years (\citealp* {vellios2022}). From a fiscal perspective, excise tax on cigarettes is a source of government revenue, which is undermined by the illicit tobacco industry.  **Blah ble** estimates the loss of income due to the illicit market to be R123123 The loss of income

# Methodology \label{methodology}
## Theoretical Model \label{Model}

This paper makes use of the demand model to estimate the effect of a relative price change, between licit and illicit cigarette price,on cigarette consumption in South Africa from January 2012 to March 2020. The relationship between cigarette consumption, price and income is well-documented (for example, see \citet{boshoff2008}; \citet{tingum2020}; \citealp{who}) and is typically expressed as:

\begin{align}
Q_{t} = f(P_{t}, Y_{t}, D_{t}) \label{eqa}
\end{align}

\ref{eqa} Where $Q_{t}$ represents cigarette consumption volume in period $t$, $P_{t}$ is the price of cigarettes adjusted for inflation, $Y_{t}$ is disposable income, and $D_{t}$ is is the dummy variable for the change in market structure that occurred in 2018. 

Two sets of prices are used in this study; the first is the legal cigarette price, which is the price of cigarettes for which duties have been paid. The second is the illicit cigarette price, which is the price of cigarettes for which duties have been paid. For this study, cigarette consumption, as represented by $Q_{t}$, is the quantity of legal cigarettes consumed (cigarettes for which duties have been paid).

## Data \label{data}

The sample period for this study runs from January 2012 to March 2020. Monthly data is used such that there are 99 observation points for each variable in the data set. One of the advantages of using monthly data rather than annual data is that it allows for more degrees of freedom. The data used in this study include figures for South Africa for the average prices of legal cigarettes, the average prices of illicit cigarettes, volume of legal cigarettes consumed, tobacco excise duties, Value Added Tax (VAT), and real disposable income. The data for legal and illicit prices, and legal cigarette consumption are provided by a prominent South African cigarette manufacturer. The other data are extracted from the South African Reserve Bank (SARB) and Statistics South Africa (StatSA).

To prepare the legal price data for analysis, the most popular price category (MPPC) is identified as the 20-cigarette pack. A weighted average of before-tax 20-pack prices is used as a base price. The excise duty per 20's pack and VAT are then added to the base price to calculate the price of legal cigarettes. The legal and illicit prices, and disposable income values, are adjusted for inflation, taking December 2016 as the base month and year, respectively. For the statistical analysis in sections \ref{Cointegrate} and \ref{vecm}, all of the variables have been transformed into log form to reduce variability. 

For the model specification, the two price series are combined to form a single price-ratio series. The price ratio is calculated as the real legal price divided by the real illicit cigarette price ($P_{legal}/P_{illicit}$). Consumer theory suggests that individuals base decisions on relative price changes rather than absolute price changes **ADD REFERENCES**. Thus, it makes intuitive sense to model the effect of a relative change in prices on the quantity of legal cigarettes

The series are plotted in levels in figure \ref{fig:plot} below. The general downwards trend of the graph depicting the real price of legal cigarettes suggests that inflation has been outstripping nominal cigarette prices. Real legal prices decreased slightly over the period 2012 - 2018. On the other hand, real prices for illicit cigarettes remained fairly constant from 2012 until late 2014. Following a small and sharp uptick, illicit prices slowly decreased from late 2014 until November 2018 when prices increase rapidly. The real price ratio is generally flat from 2012 - 2016, and then trends upwards slightly, implying that real legal prices were decreasing more slowly than real illicit prices were decreasing over the same period. There is a sudden drop in the price ratio in November 2018, reflecting the spike in real illicit prices. The graphs of the respective logged series (shown in appendix \ref{appendixLog}) illustrate similar patterns.

The plot of the real price ratio suggests that there is a structural break in the series, in late 2018. To test for this^[The test was conducted using the programming language *'R'*, using the *'strucchange'* package by \citet* {struc1} and \citet* {struc2}], the Chow structural test (based on the work by \citealp* {chow1960}) is employed for November 2018; the results are presented in table \ref{tab:struc}. The p-value is less than 0.05; thus, the null hypothesis is rejected and there is sufficient evidence to conclude that the data contains a structural breakpoint. To control for this in the model, a dummy variable is used to capture the change in the illicit price (and therefore the price ratio) that occurred in November 2018. The variable is labelled '*Structural change*' and is coded 0 for periods before November 2018, and 1 otherwise.

The graph for the quantity of legal cigarettes in figure \ref{fig:plot} shows a general downwards trend from 2012 to March 2020, and the series exhibits strong signs of seasonality. The series is decomposed and its separate components are graphed in appendix \ref{seasonGraph}, which confirm the presence of seasonality. While the series could be seasonally adjusted to account for the seasonality, this may change the outcome of conventional unit root tests, according to \citealp{season}, and lead to bias in the estimation of relationships between series (\citealp* {seasonWilliam}:110). Therefore, this study uses the unaltered data for this consumption variable in the model estimation.

The overall trends and shapes of the graphs suggest that the series are not stationary, which is formally tested below in section \ref{Cointegrate}.

```{r graphs, fig.height = 4.5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Plot for variables at levels \\label{fig:plot}"}
source("code/time_series_plots.R")
time_series_plots(log_data)
```

\newpage
\newpage
## Stationarity and Cointegration \label{Cointegrate}

To test whether the series are stationary, three tests were employed: the Augmented Dickey-Fuller^[Based on the work of \citet* {dickey}, using the '*urca*' package by \citet {urca}] (ADF) and Phillip-Perron^[Based on the work of \citet* {phillip} and using the '*tseries*' package by \citet* {tseries}] (PP) unit root tests, and the Kwiatkowski–Phillips–Schmidt–Shin^[Based on the work of \citet* {kpss}, using the '*tseries*' package by \citet* {tseries}] (KPSS) stationarity test. The results are presented in appendix \ref{appendixB}.

The results of the ADF tests in table \ref{tab:adf} show that the log of cigarette consumption, log of the real price ratio, and log of real disposable income all contain a unit root since the null hypothesis cannot be rejected, even at a 10% signifiance level. The KPSS stationarity tests (\ref{tab:kp}) confirm these results, where the null hypothesis of stationarity is rejected at a 1% level of signifiance for each series. The results of the PP tests are presented in table \ref{tab:pp}, which shows that the log of cigarette consumption does not contain a unit root at 1% significance level but that the other two series do contain unit roots.

Given that there is a structural break point in the price ratio series, standard unit root testing may not be appropriate for this series. Consequently, the Zivot-Andrews Unit Root test is used as a robustness check for the stationarity of the price ratio. The results are presented in table \ref{tab:zivot} and show that the null of a unit root cannot be rejected, at even 10% level of significance. Thus confirming the original test results.

Based on the unit root test results, the variables are assumed to be non-stationary in levels.

After first differencing each of the series, all tests indicate that the variables are stationary at a 5% level of significance. These results are presented in tables \ref{tab:adfDiff}, \ref{tab:ppDiff}, \ref{tab:kpDiff}, and \ref{tab:zivotDiff} for the ADF, PP, KPSS and Zivot-Andrews tests respectively. Given that the series are non-stationary in levels but are stationary in first differences, a cointegration approach can be used to evaluate whether a long-run relationship exists among the variables.

\
**Cointegration**

To test for conintegrating relationships, two tests were employed:
the Johansen Trace test for Cointegration, and the Johansen Maximum Eigenvalue test for Cointegration based on the work of \citealp* {johansen1990}.

The Johansen Trace Test for Cointegration indicates that there is one cointegrating vector. As shown in table \ref{trace}, the test statistic of 80.8 is larger than 60.2 (the critical value at a 1% level of significance). Thus, the null hypothesis is rejected, where the null hypothesis is that there are zero cointegrating vectors. However, the test statistic 41 is lower than 41.1 (the critical value at a 1% level of significance). This means that the null hypothesis cannot be rejected even at a 1% signicance level, where the null hypothesis is that there is 1 or fewer cointegrating vectors. 

The results of the Johansen Maximum Eigenvalue test are presented in table \ref{eigen}. Similar to the Trace test, the Maximum Eigenvalue test shows that there is a single cointegrating vector. Both cointegration tests were run with a lag length of 3, indicating that the appropriate number of lags to use in the Vector Error Correction Model is 2 lags. The presence of a cointegrating vector indicates that a long-run relationship exist among the variables. A Vector Error Correction Model is used to combine the long and short-run dynamics. This model is specified and estimated in the following section (\ref{vecm}).

## Vector Error Correction Model \label{vecm}

The equation representing the long-run equilibrium model is a double-logarithmic demand equation, given by:

\begin{align}
 lnQ_t = \alpha_0 + \alpha_1lnP_t + \alpha_2lnY_t + \alpha_3lnD_t + \mu_t \label{eq1}
\end{align}

where $i = 0,1,2,3$,\newline
$lnQ_t$ is the log of legal cigarette consumption,\newline
$lnP_t$ is the log of real price ratio for cigarettes,\newline
$lnY_t$ is the log of real disposable income,\newline
$D_t$ is the dummy variable for the structural break \newline
$\mu_t$ is the random error term, and $t$ is measured in months 

The Vector Error Correction model specification is given by:

\begin{align}
 dLnQ_t = &\beta_0 + \beta_1\sum_{i = 1}^{n-1}dLnQ_{t-i} + \beta_2\sum_{i = 1}^{n-1}dLnP_{t-i} + \beta_3\sum_{i = 1}^{n-1}dLnY_{t-i}\\ 
&+ \beta_4D_t + \lambda_iECT_{t-1} + \mu_t \label{eq2}
\end{align}

where $j - 1$ is the lag length, reduced by 1 due to the loss of a lag when differencing a Vector Autoregression (VAR),\newline
$d$ is the difference operator,\newline
$\beta_i$ represents model's short-run coefficients,\newline
$\lambda_i$ is the speed of adjustment parameter,\newline
$ECT_{t-1}$ is the error correction term, and \newline
$\mu_t$ is the random error term, and $t$ is measured in months 

# Analysis \label{analysis}

The VECM was with 1 cointegrating vector and at lag length of 2. The results are presented in table the appendix (\ref{appendixD}). The

The price ratio captures a relative change of the legal price compared to the illicit price. If the ratio shrinks, it indicates that the cost of legal cigarettes decreased relative to the cost of illicit cigarettes. There should be a negative long run relationship between the price ratio and the quantity of legal cigarettes. There should be a positive sign for the cointegrating relationship, which there is. The coefficient for the real disposable income variable should be negative for the long-run relationship (so that the reverse sign is positive); whereas the sign is positive in the Vecm results below.


The VECM was estimated at lag length of 1 with 1 cointegrating vector, and the
results are summarized in Table 2. 

The demand equation for annual cigarette consumption obtained from the VECM and 2SLS estimations is reported in Ta- bles 2. Table 3 presents the rst stage estimates of the 2SLS method. Appendix
6 shows estimated results of the demand equation using specic dummies for the
existing legislative acts in South Africa. 

The VECM uses stationary data at rst dierences and includes the lagged residuals of the long-run relationship as an explanatory variable. Coe¢ cients
from ECM represent the relationship in the short run, and the coe¢ cient of the
lagged residuals measures the speed of convergence to the long-run equilibrium.
The error correction term is the speed of adjustment in the direction of long-
term equilibrium after any deviation from the steady state. The error correction
terms in Table 2 have the correct sign and are signicant. This indicates that
per capita cigarette demand converges to steady state equilibrium at the speed
of 15% in the restricted model (see column 1) and 26% in the unrestricted model
(see column 2).
The estimate



If the real disposable income variable is excluded then the Vecm results are as follows. The coefficient for 

The previous month's deviation from the long run equilibrium is corrected at a speed of

The results of the diagnostic tests are shown in Appendix \ref{appendixE}. Table \ref{tab:serial} shows the results of the serial autocorrelation test. Table \ref{tab:arch} shows the arch results. Table \ref{tab:normal} shows the normal results.

Figure \ref{fig:residPlot} shows the residuals.

# Conclusion \label{conclusion}




